---
title: "Graphs for Report"
author: "Elena Shaw"
date: "27/06/2020"
output: pdf_document
---

```{r include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(pwr)

source("func.R")
data = read.csv('data/model_ready_data.csv')
data_pvals = read.csv('data/binom_test_data.csv')
val_data = read.csv('data/validation_ready_data.csv')
htest_results = read.csv('data/qval_CI_test_results.csv')
val_htest = read.csv('data/val_qval_CI_results.csv')
bayes_bin_results = read.csv('data/Bayesian_mcmc_scores.csv')

total_patients = 59815
```

# Refs
```{r}
data_TPs = read.csv('data/all_TPs.csv')
data_FPs = read.csv('data/potential_FP.csv')
data_COSMIC = read.csv('data/full_cosmic.csv')
ref_grp_ct = ref_overlap_ct(data, data_TPs, data_FPs, data_COSMIC)
label_ct = ref_overlap_ct(data,data_TPs, data_FPs, data_COSMIC,label=T)

val_TPs = read.csv('data/all_TPs.csv')
val_FPs = read.csv('data/potential_FP.csv')
val_COSMIC = read.csv('data/full_cosmic.csv')
val_ref_ct = ref_overlap_ct(val_data, val_TPs, val_FPs, val_COSMIC)
val_label_ct = ref_overlap_ct(val_data, val_TPs, val_FPs, val_COSMIC, label=T)

ref_labels = c("False Positive","True Positive","True Negative")
names(ref_labels) = c("FP","TP","TN")
```

## ref comparison fns
```{r}
ref_grp_perc_overlap = function(result_df) return(full_join(result_df,ref_grp_ct, by=c('ref_grp','comparison')) %>% mutate(perc_overlap = n/ref_records, ref_grp=factor(ref_grp, levels=c("FP","TP","TN"))))

label_perc_overlap = function(result_df) return(full_join(result_df,label_ct, by=c('ref_label')) %>% mutate(perc_overlap = n/ref_records, ref_label=factor(ref_label, levels=c("negatives","positives"))))

val_ref_perc_overlap = function(result_df) return(full_join(result_df,val_ref_ct, by=c('ref_grp','comparison')) %>% mutate(perc_overlap = n/ref_records))

find_divergence = function(data, x, column_name) {
  results_table = data.frame(neg_rate=numeric(length(x)), pos_rate=numeric(length(x)))
  
  for (i in 1:length(x)){
    if (i %% 50 == 0){
      print(x[i])
    }
    
    filtered_data = data %>% filter(sym(column_name) < exp(x[i]))
    if (!is.null(nrow(compare_refs(filtered_data)$div))) {
      record_overlap = compare_refs(filtered_data)$div %>%
                        mutate(ref_label = as.factor(ref_label)) %>%
                        group_by(ref_label) %>% tally()
      record_overlap_perc = label_perc_overlap(record_overlap) %>%
                              replace(is.na(.), 0)
      results_table[i,] = record_overlap_perc$perc_overlap
    }
  }
  
  return(results_table)
}
```
## GENIE patterns
```{r}
data %>% filter(Observations < 10) %>% mutate(mu_diff = obs.mu-E.mu) %>% arrange(desc(mu_diff), desc(E.mu))
```
```{r}
eg_codon = data %>% filter(Codon_Id=='ATM_G2023') %>% select(c(obs.mu,E.mu, Observations)) #TSC1_E889

x = seq(0,20,by=1)
y_E = dbinom(x,total_patients,eg_codon$E.mu)
y_O = dbinom(x,total_patients,eg_codon$obs.mu)
plot(x,y_E)#, ylim=c(0,.5),type="l")
lines(x,y_O)
```
```{r}
eg_codon = data %>% filter(Codon_Id=='ATM_G2023') %>% select(c(obs.mu,E.mu, Observations)) #TSC1_E889

x = seq(0,15,by=1)
eg_df = data.frame(x=x, E_ct=dbinom(x,total_patients,eg_codon$E.mu),
                    O_ct=dbinom(x,total_patients,eg_codon$obs.mu)) %>%
          rowwise() %>%
          mutate(min_y = min(E_ct,O_ct)) %>%
          pivot_longer(c(E_ct,O_ct), names_to = "cohort", values_to = "probability")
          
ggplot(eg_df) +
    geom_bar(data=subset(eg_df,x>=0 & x<=7 & probability<.03),aes(x=x,y=probability,fill=cohort), position="dodge", stat="identity", alpha=.5) +
    geom_bar(aes(x=x,y=probability,fill=cohort), position="dodge", stat="identity", alpha=.3) +
    geom_line(aes(x=x,y=probability,color=cohort),show.legend=F) +
    geom_vline(xintercept = total_patients * eg_codon$E.mu, 
                color = "red") +
    geom_vline(xintercept = eg_codon$Observations, 
                color = "blue") +
    theme_classic() +
    coord_cartesian(xlim=c(0,13), ylim=c(0,.25)) +
    annotate("text", x = 7, y = .23, label = "Effect size") +
    annotate("segment", x = eg_codon$Observations, 
             xend = total_patients * eg_codon$E.mu, 
             y = .22, yend = .22,
             arrow = arrow(length = unit(2, "mm"))) +
    annotate("segment", xend = eg_codon$Observations, 
             x = total_patients * eg_codon$E.mu, 
             y = .22, yend = .22,
             arrow = arrow(length = unit(2, "mm"))) +
    theme(axis.title.y = element_blank()) +
    xlab("Number of Mutations") +
    scale_fill_discrete(name = "Population", labels = c("Non-cancerous", "Cancerous"))
# ggsave("plots/Report_effect_size.pdf",width = 15, height = 7, units = "cm")
```

## Ref patterns: mutability
```{r}
pval_df = data_pvals %>% select(c(Codon_Id,obs.mu,E.mu,exact_bin))

ref_mu = compare_refs(pval_df, data_TPs, data_FPs, data_COSMIC)$all %>%
              pivot_longer(c(obs.mu,E.mu),
                           names_to = "measure",
                           values_to = "mu") %>%
              mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN")))

measure_labels = c("GENIE", "Baseline")
names(measure_labels) = c("obs.mu","E.mu")
ref_labels = c("False Positive","True Positive","True Negative")
names(ref_labels) = c("FP","TP","TN")

ggplot(ref_mu, 
       aes(x=comparison,y=mu,color=measure)) +
  geom_boxplot(show.legend=T) +
  facet_grid(~ref_grp,
             labeller = labeller(ref_grp = ref_labels),
             scale="free",
             space = "free_x") +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions"
                            )) +
  scale_color_discrete(name = "", labels = c("Baseline data", "GENIE data")) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("log(Mutability)") +
  scale_y_log10()
# ggsave("plots/Report_ref_mu.pdf")
```

## Ref patterns: (Null) Hypothesis testing
```{r}
ref_grp_perc_overlap = function(result_df) return(full_join(result_df,ref_grp_ct, by=c('ref_grp','comparison')) %>% mutate(perc_overlap = n/ref_records, ref_grp=factor(ref_grp, levels=c("FP","TP","TN"))))

# htest_results = read.csv('data/qval_CI_test_results.csv')

BF_q_discoveries = htest_results %>% filter(BF_q_reject==T) %>% mutate(disc_method = "BF_q")
BY_q_discoveries = htest_results %>% filter(BY_q_reject==T)  %>% mutate(disc_method = "BY_q")
BF_CI_discoveries = htest_results %>% filter(BF_CI_reject==T)  %>% mutate(disc_method = "BF_CI")
BY_CI_discoveries = htest_results %>% filter(BY_CI_reject==T)  %>% mutate(disc_method = "BY_CI")

all_discoveries = bind_rows(BF_q_discoveries,BY_q_discoveries,BF_CI_discoveries,BY_CI_discoveries) %>%
                    mutate(disc_method = factor(disc_method,
                                                   levels=c('BF_CI','BY_CI','BF_q','BY_q')))

all_discoveries %>% group_by(disc_method) %>% tally()

grp_results = compare_refs(all_discoveries,data_TPs, data_FPs, data_COSMIC)$all %>%
                mutate(ref_grp = as.factor(ref_grp)) %>%
                group_by(disc_method,ref_grp,comparison) %>% tally()

comp_df = ref_grp_perc_overlap(grp_results)

ggplot(comp_df, aes(x=comparison,y=perc_overlap,fill=disc_method)) +
  geom_bar(aes(fill=disc_method),
          position="dodge", stat="identity", alpha = .8) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free",space="free") +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified") +
  scale_fill_discrete(name = "", 
                      labels = c("Bonferroni Confidence Interval", 
                                 "B-Y Confidence Interval", 
                                 "Bonferroni q-value","B-Y q-value"))
# ggsave("plots/Report_HT_results.pdf")
```
### Compare pattern: HTesting
```{r}
ref_sig_val = compare_refs(htest_results,data_TPs, data_FPs, data_COSMIC)$all %>%
              mutate(lg_BF_CI = log(BF_distance),
                     lg_BY_CI = log(BY_distance)
                     ) %>%
              pivot_longer(c(BF_qval,lg_BF_CI,BY_qval, lg_BY_CI), 
                           names_to = "disc_method",
                           values_to = "disc_val") %>%
              mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN")),
                     disc_method = factor(disc_method,
                                          levels=c('lg_BF_CI','lg_BY_CI','BF_qval','BY_qval')),
                     measure = ifelse(disc_method=="BF_qval"|
                                      disc_method=="BY_qval",
                                      "p-value","Confidence Interval"))

ggplot(ref_sig_val) +
  geom_boxplot(aes(x=comparison,y=disc_val,color=disc_method)) +
  facet_grid(measure~ref_grp,
             labeller = labeller(ref_grp = ref_labels),
             scale="free",
             space = "free_x") +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions"
                            )) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position="top"
        ) +
  ylab("       adjusted p-value            log(distance above C.I.)") +
  scale_color_discrete(name = "", 
                      labels = c("Bonferroni Confidence Interval", 
                                 "B-Y Confidence Interval", 
                                 "Bonferroni q-value","B-Y q-value"))
   
# ggsave("plots/Report_ref_qCI.pdf", height = 15, units="cm")
```
## Discriminatory power: alpha
```{r}
alpha = seq(-800,0,by=.1)

alpha_table = find_divergence(htest_results, alpha, 'BF_qval')
alpha_table = find_divergence(htest_results, alpha, 'BY_qval')

alpha_results = bind_cols(alpha_table,list(a=alpha))
# write.csv(alpha_results,'data/divergence_alpha_BF.csv',row.names=F)
# write.csv(alpha_results,'data/divergence_alpha_BY.csv',row.names=F)
# alpha_results = read.csv('data/divergence_alpha_BF.csv')
alpha_results = read.csv('data/divergence_alpha_BY.csv')

a_divergence = alpha_results %>% mutate(rate_diff = abs(pos_rate-neg_rate)) %>% arrange(desc(rate_diff)) %>% top_n(1)

alpha_roc =  alpha_results%>%
              filter(neg_rate>0, pos_rate>0) %>%
              pivot_longer(c(pos_rate,neg_rate),
                              names_to = "label",
                              values_to = "rate"
                              ) %>%
              mutate(label = factor(label,levels=c("pos_rate","neg_rate")))

ggplot(alpha_roc, aes(x=a,y=rate,color=label)) +
  geom_line() +
  geom_vline(xintercept = a_divergence$a, linetype="dashed")+
  scale_color_discrete(name = "",
                      labels = c("Potential Positives","Potential Negatives")) +
  theme(legend.position="top") +
  coord_cartesian(xlim = c(-50,0), ylim=c(0,1)) +
  annotate("text", x = a_divergence$a-7, y = .45, 
           label = paste0("log(adjusted p) = ", round(a_divergence$a,3))) +
  annotate("text", x = a_divergence$a-6, y = .4, 
           label = paste0("separation = ", round(a_divergence$rate_diff,3))) +
  xlab("log(B-Y adjusted p-value)") + ylab("Proportion")
  # xlab("log(Bonferroni adjusted p-value)") + ylab("Proportion")

# ggsave('plots/Report_div_a_BF.pdf', height=11,units="cm")
# ggsave('plots/Report_div_a_BY.pdf', height=11,units="cm")

```

## Ref patterns: alpha v. beta
```{r}
std_power = function(h) return(pwr.p.test(h=h, n=total_patients, 
                                          sig.level=.01,
                                          alternative="greater")$power)

pwr_df = htest_results %>%
          select(c(Codon_Id,obs.mu,E.mu,BF_qval,BY_qval)) %>%
          mutate(effect_size_h = 2*(asin(sqrt(obs.mu)) - asin(sqrt(E.mu))),
                 power = std_power(effect_size_h),
                  beta = 1- power
          )

ref_pwr = compare_refs(pwr_df, data_TPs, data_FPs, data_COSMIC)$all  %>%
            mutate(ref_grp = factor(ref_grp,levels=c("FP","TP","TN")))

comparison_labels = c("cBio_TP" = "Chang True Positives", 
                      "Chang_TP" = "Chang Predictions",
                      "consensus" = "Consensus Positives",
                      "Brown_TP" = 'Brown "Driver"',
                      "Brown_FP" = 'Brown "Passenger"',
                      "Mutagene_TP" = 'Li "Non-netural"',
                      "Mutagene_FP" = 'Li "Netural"',
                      "UNK" = "No mentions"
                      )
ref_labels = c("False Positive","True Positive","True Negative")
names(ref_labels) = c("FP","TP","TN")

ggplot(ref_pwr, aes(x=comparison,y=log(beta),color=comparison)) +
  geom_boxplot(show.legend = F)+
  facet_grid(~ref_grp,
             labeller = labeller(ref_grp = ref_labels),
             scale="free",
             space = "free_x") +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions"
                            )) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        ) +
  ylab("log(Type II error)")
# ggsave("plots/Report_ref_power.pdf")
```
## Discriminatory power: beta
```{r}
beta = seq(-40,0,by=.01)
beta_table = find_divergence(pwr_df, beta, 'beta')
beta_results = bind_cols(beta_table,list(b=beta))

# write.csv(alpha_results,'data/divergence_beta.csv',row.names=F)
beta_results = read.csv('data/divergence_beta.csv')
b_divergence = beta_results %>% 
                  mutate(rate_diff = abs(pos_rate-neg_rate)) %>% 
                  arrange(desc(rate_diff)) %>% 
                  top_n(1)

beta_roc =  beta_results%>%
          pivot_longer(c(pos_rate,neg_rate),
                          names_to = "label",
                          values_to = "rate"
                          ) %>%
          mutate(label = factor(label,levels=c("pos_rate","neg_rate")))

ggplot(beta_roc, aes(x=b,y=rate,color=label)) +
  geom_line() +
  geom_vline(xintercept = b_divergence$b, linetype="dashed")+
  scale_color_discrete(name = "",
                      labels = c("Potential Positives","Potential Negatives")) +
  theme(legend.position="top") + xlab("log(Type II error)") +
  ylab("Proportion") +
  coord_cartesian(xlim = c(-40,0), ylim=c(0,1)) +
  annotate("text", x = b_divergence$b-4.5, y = .45, 
           label = paste0("log(error) = ", b_divergence$b)) +
  annotate("text", x = b_divergence$b-4.5, y = .4, 
           label = paste0("separation = ", round(b_divergence$rate_diff,3)))
# ggsave('plots/Report_div_b.pdf')
```

```{r}
high_power_BF = pwr_df %>% filter(beta < max(b_divergence$b), BF_qval<a_divergence$a)
high_power_BY = pwr_df %>% filter(beta < max(b_divergence$b), BY_qval<a_divergence$a)
       
high_power_BF_overlap = compare_refs(high_power_BF,data_TPs, data_FPs, data_COSMIC)$all %>%
                      group_by(ref_grp,comparison) %>%
                      tally()
high_power_BY_overlap = compare_refs(high_power_BY,data_TPs, data_FPs, data_COSMIC)$all %>%
                      group_by(ref_grp,comparison) %>%
                      tally()

high_power_BF_perc = ref_grp_perc_overlap(high_power_BF_overlap) %>% 
                      replace(is.na(.), 0) %>% 
                      mutate(disc_method="BF")
high_power_BY_perc = ref_grp_perc_overlap(high_power_BY_overlap) %>% 
                      replace(is.na(.), 0) %>% 
                      mutate(disc_method="BY")

just_a_perc = comp_df %>% filter(disc_method == "BF_q") #| disc_method == "BY_q")

high_power_perc = bind_rows(high_power_BF_perc,just_a_perc) %>%
                    mutate(disc_method = factor(disc_method,
                                                levels=c("BF_q","BF"))) #,"BY_q","BY")))

ggplot(high_power_perc, aes(x=comparison,y=perc_overlap)) +
  geom_bar(aes(fill=disc_method),
          position="dodge", stat="identity", alpha = .8) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free", space="free"
             ) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified") +
  scale_fill_discrete(name = "", 
                      labels = c("Type I error only (.01)", 
                                 "Type I & Type II errors")) 

# ggsave("plots/Report_aVb_perc_BF.pdf")

```
```{r}
pixel_size = c(50,.01)

aXb_ref = ref_pwr %>% pivot_longer(c(BF_qval,BY_qval),
                         names_to = "metric",
                         values_to = "alpha") %>%
          filter(beta>0)

ggplot(aXb_ref, aes(x=log(alpha), y=log(beta), color=metric)) +
  geom_point(alpha=.3) +
  facet_wrap(~ref_grp,
             labeller = labeller(ref_grp = ref_labels, comparison = comparison_labels)) +
  xlab("log(Type I error)") + ylab("log(Type II error)") +
  theme(legend.position="top") +
  scale_color_discrete(name = "", 
                      labels = c("Bonferroni adjusted p","B-Y adjusted p"))

# ggsave("plots/Report_ref_aVb.pdf")
```


## Check results with validation set
note: even at .01 q-val threshold, we only get 87 results. If we set beta at .2, down to 37 results. At max divergence threshold, 0 records.
```{r}
# val_htest = read.csv('data/qval_validation_results.csv')
# 
# val_sig_results = val_htest #%>%
#                     # filter(BF_qval<a_divergence$a) #, BY_qval<a_divergence$a)
# 
# val_pwr = val_sig_results %>%
#           select(c(Codon_Id,obs.mu,E.mu,BF_qval,BY_qval)) %>%
#           mutate(effect_size_h = 2*(asin(sqrt(obs.mu)) - asin(sqrt(E.mu))),
#                  power = std_power(effect_size_h),
#                   beta = 1- power
#           ) %>% filter(beta < max(b_divergence$b))
# 
# val_ht_results = compare_refs(val_pwr,val_TPs, val_FPs, val_COSMIC)$val %>%
#                 mutate(ref_grp = as.factor(ref_grp)) %>%
#                 group_by(ref_grp,comparison) %>% tally()
# 
# val_ht_overlap = val_ref_perc_overlap(val_ht_results) %>%
#                   replace(is.na(.), 0)
# 
# ggplot(val_ht_overlap, aes(x=comparison,y=perc_overlap,fill=comparison,label=n)) +
#   geom_col(show.legend = F, alpha=.8) +
#   geom_text(position = position_dodge(width = 0.9),
#             vjust = 3) +
#   scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
#                             "Chang_TP" = "Chang Predictions",
#                             "consensus" = "Consensus Positives",
#                             "Brown_TP" = 'Brown "Driver"',
#                             "Brown_FP" = 'Brown "Passenger"',
#                             "Mutagene_TP" = 'Li "Non-netural"',
#                             "Mutagene_FP" = 'Li "Netural"',
#                             "UNK" = "No mentions")) +
#   facet_grid(~ref_grp, 
#              labeller = labeller(ref_grp = ref_labels),
#              scale = "free", space="free"
#              ) +
#   theme(axis.text.x = element_text(angle=45,hjust=1),
#         axis.title.x = element_blank(),
#         legend.position="top"
#         ) +
#   ylab("Proportion Identified")
```

## Bayesian binom test
```{r}
# std_power = function(h) return(pwr.p.test(h=h, n=total_patients, 
#                                           sig.level=.01,
#                                           alternative="greater")$power)

bayes_discoveries = bayes_bin_results %>% 
                      mutate(bayes_p = exp(log_score)) %>%
                      filter(bayes_p < .01)
       
bayes_disc_overlap = compare_refs(bayes_discoveries)$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()

bayes_perc_df = ref_grp_perc_overlap(bayes_disc_overlap)


new_comp_df = bind_rows(comp_df,bayes_perc_df %>% 
                          mutate(disc_method = "Bayes")) %>%
                  mutate(disc_method = factor(disc_method,
                                          levels=c('BF_CI','BY_CI','BF_q','BY_q','Bayes'))
                        )

ggplot(new_comp_df, aes(x=comparison,y=perc_overlap)) +
  geom_bar(aes(fill=disc_method),
          position="dodge", stat="identity", alpha = .8) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free", space="free") +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified") +
  scale_fill_discrete(name = "", 
                      labels = c("Bonferroni C.I.", "B-Y C.I.", 
                                 "Bonferroni q-value","B-Y q-value",
                                 "Bayes p-value"))
# ggsave('plots/Report_bayes_test_ref.pdf',height=12,units = "cm")
```

## Cross-sectional metrics
```{r}
val_perc_label = function(result_df) return(full_join(result_df,val_label_ct, by=c('ref_label')) %>% mutate(perc_overlap = n/ref_records))

val_risk_ES_data = val_data %>%
                    mutate(risk_ratio = obs.mu/E.mu,
                           effect_size_h = 2*(asin(sqrt(obs.mu)) - asin(sqrt(E.mu))))

filter_rate = function(crit_risk,crit_ES) {
  filtered_data = val_risk_ES_data %>% filter(risk_ratio > crit_risk, effect_size_h > -exp(crit_ES))
  record_overlap = compare_vals(filtered_data)$val %>%
                        mutate(ref_label = as.factor(ref_label)) %>%
                        group_by(ref_label) %>% tally()
  record_overlap_perc = val_perc_label(record_overlap) %>%
                          replace(is.na(.), 0)
                          
  return(record_overlap_perc)
}
```

## Effect Size
```{r}
ES_df = data_pvals %>%
          mutate(effect_size_h = 2*(asin(sqrt(obs.mu)) - asin(sqrt(E.mu))))

ref_ES = compare_refs(ES_df)$all %>%
            mutate(ref_grp = factor(ref_grp,levels=c("FP","TP","TN")))

ggplot(ref_ES, aes(x=comparison,y=log(effect_size_h),color=comparison)) +
  geom_boxplot(show.legend = F)+
  facet_grid(~ref_grp,
             labeller = labeller(ref_grp = ref_labels),
             scale="free",
             space = "free_x") +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions"
                            )) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        ) 

# ggsave('plots/Report_ref_effect_size.pdf')

high_ES = ES_df %>% filter(effect_size_h>exp(-4.5)) %>% select(Codon_Id) %>% unique()

sig_ES_overlap = compare_refs(high_ES,data_TPs, data_FPs, data_COSMIC)$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()

sig_ES_perc = ref_grp_perc_overlap(sig_ES_overlap)
```
## Discriminatory power: Effect Size
```{r}
es1 = seq(-10,0,by=.01)
es1_table = find_divergence(ES_df, es1, 'effect_size_h')
es1_results = bind_cols(es1_table,list(es=es1))

# write.csv(es1_results,'data/divergence_es1_data.csv')
es1_results = read.csv('data/divergence_es1_data.csv')
es1_divergence = es1_results %>% mutate(rate_diff = abs(pos_rate-neg_rate)) %>% arrange(desc(rate_diff)) %>% top_n(1)

es1_roc =  es1_results%>%
          pivot_longer(c(pos_rate,neg_rate),
                          names_to = "label",
                          values_to = "rate"
                          ) %>%
          mutate(label = factor(label,levels=c("pos_rate","neg_rate")))

ggplot(es1_roc, aes(x=es,y=rate,color=label)) +
  geom_line() +
  geom_vline(xintercept = es1_divergence$es, linetype="dashed")+
  scale_color_discrete(name = "",
                      labels = c("Potential Positives","Potential Negatives")) +
  theme(legend.position="top") + xlab("log(Effect Size)") +
  ylab("Proportion") +
  coord_cartesian(xlim = c(-10,0), ylim=c(0,1)) +
  annotate("text", x = es1_divergence$es+1.3, y = .95, 
           label = paste0("log(effect size) = ", es1_divergence$es)) +
  annotate("text", x = es1_divergence$es+1.1, y = .9, 
           label = paste0("separation = ", round(es1_divergence$rate_diff,3)))
ggsave("plots/Report_div_es1.pdf") #,width = 15, height = 7, units = "cm")
```

## validation
```{r}
val_high_ES = val_risk_ES_data %>%
                    filter(effect_size_h > exp(es1_divergence$es))

val_high_ES_results = compare_refs(val_high_ES,val_TPs, val_FPs, val_COSMIC)$all %>%
                mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN"))) %>%
                group_by(ref_grp,comparison) %>% tally()

val_ES_eval = val_ref_perc_overlap(val_high_ES_results) %>%
                  replace(is.na(.), 0)

ggplot(val_ES_eval %>%
                mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN"))),
       aes(x=comparison,y=perc_overlap,fill=comparison,label=n)) +
  geom_col(show.legend = F, alpha=.8) +
  geom_text(position = position_dodge(width = 0.9),
            vjust = -.2) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free", space="free"
             ) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  coord_cartesian(ylim=c(0,.08)) +
  ylab("Proportion Identified")
# ggsave("plots/Report_val_es_eval.pdf")
```

## Risk
## Ref patterns: Risk
```{r}
# Risk Ratio:
# "the quotient between the risks, resp. probabilities for incidences in two different groups. The risk is computed by dividing the number of incidences by the total number in each group and building the ratio between the groups."

# Odds Ratio:
# "If for example 10 persons die in a group and 90 survive, than the odds in the groups would be 10/90, whereas the risk would be 10/(90+10). The odds ratio is the quotient between the odds of the two groups."

# Risk Difference:
# "the difference between two risks. Compared to the ratios, the risks are not divided but subtracted from each other. For the computation of Risk Differences, only the raw data is used, even when calculating variance and standard error. The measure has a disadvantage: It is highly influenced by changes in base rates."
pval_df = data_pvals %>% select(c(Codon_Id,obs.mu,E.mu,exact_bin))

risks_df = pval_df %>%
            mutate(risk_ratio = obs.mu/E.mu,
                   odds_ratio = (obs.mu/(1-obs.mu))/(E.mu/(1-E.mu)),
                   risk_diff = (obs.mu - E.mu)/total_patients
                   )

all_riskXq = compare_refs(risks_df)$all %>%
              pivot_longer(c(risk_ratio,odds_ratio,risk_diff),
                           names_to = "measure",
                           values_to = "risk") %>%
              mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN")))

measure_labels = c("Risk Ratio", "Risk Difference")
names(measure_labels) = c("risk_diff","risk_ratio")
ref_labels = c("False Positive","True Positive","True Negative")
names(ref_labels) = c("FP","TP","TN")

ggplot(all_riskXq %>% filter(measure == "risk_ratio"), 
       aes(x=comparison,y=log(risk),color=comparison)) +
  geom_boxplot(show.legend=F) +
  facet_grid(.~ref_grp,
             labeller = labeller(measure = measure_labels, ref_grp = ref_labels),
             scale="free",
             space = "free_x") +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions"
                            )) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        ) +
  ylab("log(Risk Ratio)")
# ggsave("plots/Report_ref_risk_ratio.pdf")
```
```{r}
sig_risk = all_discoveries %>% mutate(risk_ratio = obs.mu/E.mu) %>% filter(risk_ratio > exp(3)) %>% select(Codon_Id) %>% unique()

sig_risk_overlap = compare_refs(sig_risk)$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()

sig_risk_perc = ref_grp_perc_overlap(sig_risk_overlap)

all_risk = ref_sig_val %>%
            mutate(risk_ratio = obs.mu/E.mu)

all_risk_labeled = compare_refs(all_risk)$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()
```
## Discriminatory power: risk
```{r}
rr1 = seq(-1,12,by=.01)
rr1_table = find_divergence(risks_df, rr1, 'risk_ratio')
rr1_results = bind_cols(rr1_table,list(rr=rr1))

# write.csv(alpha_results,'data/divergence_rr1_data.csv')
rr1_results = read.csv('data/divergence_rr1_data.csv')
rr1_divergence = rr1_results %>% mutate(rate_diff = abs(pos_rate-neg_rate)) %>% arrange(desc(rate_diff)) %>% top_n(1)

rr1_roc =  rr1_results%>%
          pivot_longer(c(pos_rate,neg_rate),
                          names_to = "label",
                          values_to = "rate"
                          ) %>%
          mutate(label = factor(label,levels=c("pos_rate","neg_rate")))

ggplot(rr1_roc, aes(x=rr,y=rate,color=label)) +
  geom_line() +
  geom_vline(xintercept = rr1_divergence$rr, linetype="dashed")+
  scale_color_discrete(name = "",
                      labels = c("Potential Positives","Potential Negatives")) +
  theme(legend.position="top") + xlab("log(Risk Ratio)") +
  ylab("Proportion") +
  coord_cartesian(xlim = c(-1,10), ylim=c(0,1)) +
  annotate("text", x = rr1_divergence$rr+1.5, y = .85, 
           label = paste0("log(Risk Ratio) = ", rr1_divergence$rr)) +
  annotate("text", x = rr1_divergence$rr+1.3, y = .8, 
           label = paste0("separation = ", round(rr1_divergence$rate_diff,3)))
  
# ggsave("plots/Report_div_rr1.pdf")
```

## Check risks with validation set
```{r}
val_ref_perc_overlap = function(result_df) return(full_join(result_df,val_ref_ct, by=c('ref_grp','comparison')) %>% mutate(perc_overlap = n/ref_records))

val_high_risk = val_risk_ES_data %>%
                    filter(risk_ratio > exp(rr1_divergence$rr))

val_high_risk_results = compare_refs(val_high_risk, val_TPs, val_FPs, val_COSMIC)$all %>%
                mutate(ref_grp = as.factor(ref_grp)) %>%
                group_by(ref_grp,comparison) %>% tally()

val_high_risk_eval = val_ref_perc_overlap(val_high_risk_results) %>%
                  replace(is.na(.), 0)

ggplot(val_high_risk_eval %>% mutate(ref_grp = factor(ref_grp,
                                         levels=c("FP","TP","TN"))),
       aes(x=comparison,y=perc_overlap,fill=comparison,label=n)) +
  geom_col(show.legend = F, alpha=.8) +
  geom_text(position = position_dodge(width = 0.9),
            vjust = -0.2) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free", space="free"
             ) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified")

# ggsave("plots/Report_val_risk_eval.pdf")
```
## High Risk x Effect Size
```{r}
risk_ES_data = pwr_df %>% mutate(risk_ratio = obs.mu/E.mu) 

riskXes_df = risk_ES_data %>% filter(risk_ratio > exp(rr1_divergence$rr), effect_size_h > exp(es1_divergence$es)) #inner_join(sig_risk,high_ES,by='Codon_Id') %>% select(Codon_Id) %>% unique()

high_risk_df = risk_ES_data %>% filter(risk_ratio > exp(rr1_divergence$rr))
high_risk_labeled = compare_refs(high_risk_df,'data/all_TPs.csv','data/potential_FP.csv',
                          'data/full_cosmic.csv')$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()
sig_risk_perc = ref_grp_perc_overlap(high_risk_labeled) %>%
                mutate(disc_method = "Risk")
high_ES_df = risk_ES_data %>% filter(effect_size_h > exp(es1_divergence$es))
high_ES_labeled = compare_refs(high_ES_df,'data/all_TPs.csv','data/potential_FP.csv',
                          'data/full_cosmic.csv')$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()
sig_ES_perc = ref_grp_perc_overlap(high_ES_labeled) %>%
                mutate(disc_method = "ES")

riskXes_labeled = compare_refs(riskXes_df,'data/all_TPs.csv','data/potential_FP.csv',
                          'data/full_cosmic.csv')$all %>%
                      group_by(comparison,ref_grp) %>%
                      tally()

# riskXes_perc = ref_grp_perc_overlap(riskXes_labeled) %>%
#                 mutate(disc_method = "RxES")


BFq_thres = all_discoveries %>% filter(BF_qval < exp(-1))
BYq_thres = all_discoveries %>% filter(BY_qval < exp(-7.3))

BF_results = compare_refs(BFq_thres,'data/all_TPs.csv','data/potential_FP.csv',
                          'data/full_cosmic.csv')$all %>%
                mutate(ref_grp = as.factor(ref_grp)) %>%
                group_by(disc_method,ref_grp,comparison) %>% tally()
BY_results = compare_refs(BFq_thres,'data/all_TPs.csv','data/potential_FP.csv',
                          'data/full_cosmic.csv')$all %>%
                mutate(ref_grp = as.factor(ref_grp)) %>%
                group_by(disc_method,ref_grp,comparison) %>% tally()

BF_perc = ref_grp_perc_overlap(BF_results) %>% mutate(disc_method="BF_q")
BY_perc = ref_grp_perc_overlap(BY_results) %>% mutate(disc_method="BY_q")

all_metrics = bind_rows(BF_perc,BY_perc,
                        sig_risk_perc, sig_ES_perc
                                 ) %>%
                        mutate(disc_method = factor(disc_method,
                                               levels=c("ES","Risk","BF_q","BY_q")))


ggplot(all_metrics, aes(x=comparison,y=perc_overlap)) +
  geom_bar(aes(fill=disc_method),
          position="dodge", stat="identity", alpha = .8) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free",space="free") +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified") +
  scale_fill_discrete(name = "", 
                      labels = c("Effect Size", "Risk ratio", 
                                 "Bonferroni q-value",
                                 "B-Y q-value"))
ggsave('plots/Report_CS_results.pdf')
```
## Check high Risk x Effect Size on validation
```{r}
risks_ES_data = pval_df %>%
                  mutate(risk_ratio = obs.mu/E.mu,
                         )

val_rXES_df = val_risk_ES_data %>%
                    filter(risk_ratio > exp(3), effect_size_h > exp(-4.5))

val_rXES_results = compare_refs(val_rXES_df, val_TPs, val_FPs, val_COSMIC)$all %>%
                        mutate(ref_grp = as.factor(ref_grp)) %>%
                        group_by(ref_grp,comparison) %>% tally()

val_rXES_eval = val_ref_perc_overlap(val_rXES_results) %>%
                  replace(is.na(.), 0)



ggplot(val_risk_eval, aes(x=comparison,y=perc_overlap,fill=comparison,label=n)) +
  geom_col(show.legend = F, alpha=.8) +
  geom_text(position = position_dodge(width = 0.9),
            vjust = 3) +
  scale_x_discrete(labels=c("cBio_TP" = "Chang True Positives", 
                            "Chang_TP" = "Chang Predictions",
                            "consensus" = "Consensus Positives",
                            "Brown_TP" = 'Brown "Driver"',
                            "Brown_FP" = 'Brown "Passenger"',
                            "Mutagene_TP" = 'Li "Non-netural"',
                            "Mutagene_FP" = 'Li "Netural"',
                            "UNK" = "No mentions")) +
  facet_grid(~ref_grp, 
             labeller = labeller(ref_grp = ref_labels),
             scale = "free", space="free"
             ) +
  theme(axis.text.x = element_text(angle=45,hjust=1),
        axis.title.x = element_blank(),
        legend.position="top"
        ) +
  ylab("Proportion Identified")
```
